Multi-Core Model Checking and Maximum Satisfiability Applied to Hardware-Software Partitioning

### 2015-12-02 Docs/IJES/journal.tex
brand new complete version of IJES paper

\authorA{Alessandro Trindade, Hussama Ismail}
\affA{Federal University of Amazonas \\
E-mail: \{alessandro.b.trindade, hussamaismail\}@gmail.com}

\authorB{Edilson Galv\~ao, Renato Degelo}
\affB{Federal University of Amazonas \\
E-mail: \{rdegelo, esj.galvao\}@gmail.com}

\authorC{Helder Silva, and Lucas Cordeiro}
\affC{Federal University of Amazonas \\
E-mail: prhsilva2012@gmail.com, lucascordeiro@ufam.edu.br}

\begin{abstract}
We present an alternative approach to solve the hardware and software partitioning problem, which uses Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) in conjunction with a multi-core support using Open Multi-Processing. In a nutshell, the multi-core SMT-based BMC approach allows initializing many verification instances based on the number of available processing cores. Each instance checks for a different optimum value until the optimization problem is satisfied. We implement our algorithms on top of the Efficient SMT-Based Context-Bounded Model Checker (ESBMC) and also integrate the maximum satisfiability solver $\nu$Z tool into ESBMC. We compare all our approaches to another state-of-the-art optimization tool (Matlab). Experimental results show that there is no single optimization tool to solve all hardware-software partitioning benchmarks; however, Matlab and ESBMC-$\nu$Z are the most efficient ones to solve hardware-software partitioning problems, although multi-core ESBMC had a significant performance improvement in particular cases.
\end{abstract}

\KEYWORD{hardware-software co-design, hardware-software partitioning, optimization, model checking, multi-core, OpenMP.}

\REF{\textcolor{red}{to this paper should be made as follows: Rodr\'{\i}guez
Bol\'{\i}var, M.P. and Sen\'{e}s Garc\'{\i}a, B. (xxxx) `The
corporate environmental disclosures on the internet: the case of
IBEX 35 Spanish companies', {\it International Journal of Metadata,
Semantics and Ontologies}, Vol. x, No. x, pp.xxx\textendash xxx.}}

\begin{bio}
\textcolor{red}{
Manuel Pedro Rodr\'iguez Bol\'ivar received his PhD in Accounting at
the University of Granada. He is a Lecturer at the Department of
Accounting and Finance, University of Granada. His research
interests include issues related to conceptual frameworks of
accounting, diffusion of financial information on Internet, Balanced
Scorecard applications and environmental accounting. He is author of
a great deal of research studies published at national and
international journals, conference proceedings as well as book
chapters, one of which has been edited by Kluwer Academic
Publishers.\vs{9}}

\textcolor{red}{\noindent Bel\'en Sen\'es Garc\'ia received her PhD in Accounting at
the University of Granada. She is a Lecturer at the Department of
Accounting and Finance, University of Granada. Her research
interests are related to cultural, institutional and historic
accounting and in environmental accounting. She has published
research papers at national and international journals, conference
proceedings as well as chapters of books.\vs{8}}

\textcolor{red}{\noindent Both authors have published a book about environmental
accounting edited by the Institute of Accounting and Auditing,
Ministry of Economic Affairs, in Spain in October 2003.}
\end{bio}

@@

Starting at the 1990s, intensive research was performed in HW-SW partitioning, and several approaches proposed, as shown in~\cite{Arato2003} and~\cite{Mann2007}. In~\cite{Trindade2015,Trindade2016} was shown that it is possible to use Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT), implemented in a tool called Efficient SMT-Based Context-Bounded Model Checker (ESBMC), in order to perform HW-SW partitioning in embedded systems. The present work extends those studies since there is a substantial improvement in terms of the SMT-based verification method, which has been extended with a multi-core parallel- and binary-search approaches, as well as, the integration of the Maximum SMT (MaxSMT) solver $\nu$Z into ESBMC, which is a state-of-art optimization tool based on SMT~\cite{Bjorner2015}.

Here, we exploit the availability of multi-core processors; in particular, SMT-based verification methods are applied to the HW-SW partition problem in three different ways using a multi-core ESBMC approach with OpenMP~\cite{Tang2009}: ESBMC-SS using a sequential-search (SS), ESBMC-PS using a parallel-search (PS), and ESBMC-PB using a binary-search (BS). Experimental results are compared to ILP (integral linear programming), GA (generic algorithms) in a multi-core version, and also to $\nu$Z, which supports only a single-core approach~\cite{Bjorner2015}. The ILP and GA algorithms are implemented with the optimization toolbox of Matlab~\cite{TheMathWorks2013}, while $\nu$Z is a built-in tool to the SMT solver Z3. All multi-core ESBMC approaches, together with $\nu$Z, are implemented with the ESBMC tool~\cite{Cordeiro2012}.

@@

Optimization is the act of obtaining the best result ({\it i.e.}, the optimal solution) under given circumstances~\cite{Rao2009}. There is no single method available for efficiently solving all optimization problems~\cite{Rao2009}. The most well-known technique is linear programming, which is an method applicable for the solution of problems in which the objective function and the constraints appear as linear functions of the decision variables. A particular case of linear programming is ILP, in which the variables can assume just integer values. Eq.~\ref{linear-programming-problem} shows a typical linear programming problem, where $A$ and $b$ are vectors or matrixes that describe the constraints

@@

\vspace{2 mm}
In some cases, the time to find a solution using ILP is impractical. Even with the use of powerful computers, a problem can take hours before an optimal solution is reached. If the optimization problem is complex, some heuristics can be used to solve the same problem faster, {\it e.g.}, those used in the GA~\cite{Rao2009}. The only drawback is that the found solution may not be the global minimum or maximum. Alternatively, tools such as ESBMC and $\nu$Z can be used to solve optimization problems so that the global minimum or maximum solution is found. The following sections describe the main features of ESBMC and $\nu$Z tools.

@@

Among the recent model checking techniques, there is one that combines model checking with satisﬁability solving. This technique, known as bounded model checking (BMC), does a very fast exploration of the state space, and for some types of problems, it offers large performance improvements over previous approaches, as shown in~\cite{Biere2009}. In particular, BMC based on Boolean Satisfiability (SAT) has been introduced as a complementary technique to binary decision diagrams for alleviating the state explosion problem~\cite{Clarke2001}.

The basic idea of BMC is to check the negation of a given property at a given depth: given a transition system $M$, a property $\phi$, and a bound $k$, BMC unrolls the system $k$ times and translates it into a verification condition (VC) $\psi$  such that $\psi$ is satisfiable if and only if $\phi$ has a counterexample of depth $k$ or less~\cite{Biere2009}. To cope with increasing software complexity, SMT solvers can be used as back-ends for solving the generated VCs, as shown in~\cite{Cordeiro2012},~\cite{Armando2009},~\cite{Ganai2006}.

In this study, ESBMC has been used as a BMC tool to solve HW-SW partitioning problems~\cite{Cordeiro2012}. 
In particular, there are two directives in ESBMC that can be used to guide it to solve an optimization problem: ASSUME and ASSERT. The directive ASSUME is responsible for ensuring the compliance of constraints (software costs), and the directive ASSERT controls the halt condition (minimum hardware cost). Then, with some C/C++ code, it is possible to guide ESBMC to solve optimization problems.

@@

Fig.~\ref{ESBMC-Architecture} shows the current ESBMC architecture, which consists of the C/C$++$ parser, GOTO Program, GOTO Symex, and SMT solver ~\cite{Ramalho2013}. In particular, ESBMC compiles C/C$++$ code into equivalent GOTO-programs ({\it i.e.}, control-flow graphs) using a gcc-compliant style. GOTO-programs can then be processed by the symbolic execution engine, called GOTO Symex, where two recursive functions compute the constraints ($C$) and properties ($P$); finally, it generates two sets of equations ({\it i.e.},\:$C \land \neg P$), which are checked for satisfiability by an SMT solver.

The main factor for ESBMC to use only a single-core relies on its back-end ({\it i.e.}, SMT Solver). Currently, the SMT solvers supported by ESBMC are: Z3~\cite{DeMoura2008}, Boolector~\cite{Brummayer2009}, MathSAT~\cite{Barrett2011}, CVC4~\cite{Bozzano2005}, and Yices~\cite{Dutertre2014}. Most of them do provide neither multi-threaded support nor a parallel version to solve the generated SMT equations.

@@

The OpenMP is a set of directives for parallel programming that augments C/C$++$ and Fortran languages~\cite{OpenMP1998}. OpenMP supports most processor architectures and operating systems, {\it e.g.}, Solaris, AIX, HP-UX, Linux, Mac OS X, and Windows. OpenMP uses a portable and very robust model to facilitate the development of parallel applications for a variety of platforms.

In particular, OpenMP uses the \textit{fork-join} model of parallel execution~\cite{OpenMP1998}. The main thread executes the sequential parts of the program; if a parallel region is encountered, then it forks a team of worker threads. After the parallel region finishes ({\it i.e.}, the API waits until all threads terminate), then the main procedure returns to the single-threaded execution mode~\cite{Wu2014}.

@@

In this study, the SMT solver Z3 is used to check for the satisfiability of formulas generated from the HW-SW partitioning problem~\cite{Bjorner2014}. In particular, we exploit the use of MaxSMT solver $\nu$Z, which is implemented on top of the SMT solver Z3, in order to solve optimization problems; $\nu$Z base function is to optimize objective functions, which formulate optimized criteria, within the logical context of constraints.~$\nu$Z also includes an incremental version of the Maximum Resiliency (MaxRes)~\cite{Federica2008}, in order to achieve Maximum Satisfiability (MaxSAT)~\cite{NarodytskaN} and a Simplex to solve numbers without defined patterns.

In $\nu$Z, MaxSAT is responsible for the restrictions, while OptSMT optimizes linear arithmetic objectives~\cite{Bjorner2015}. In summary, $\nu$Z provides three main functions that extend Z3 for solving optimization problems, which are: \textit{maximize}, \textit{minimize}, and \textit{assert-soft}.

@@

\vspace{2 mm}
Fig.~\ref{vZ-Architecture} shows the $\nu$Z architecture. Initially, the SMT formula with objectives is converted to $0-1$ constraints, which leads to a Pseudo-Boolean Optimization (PBO)~\cite{Barth1995,Vasco2005}. If there are many objective functions, $\nu$Z invokes OptSAT for arithmetic or MaxSAT for soft constraints. For constraints using real values, $\nu$Z combines linear arithmetic objectives and uses only one instance of OptSMT. When ``soft constrains'' is used in the mode `` lexicographic'', $\nu$Z invokes MaxSAT using multiple calls for its engine.

@@

Z3 is available for platforms in C, C$++$, Java, .NET, and Python; it is possible to download Z3 with $\nu$Z from its github repository~\cite{Z3API}. In this work, the python API is used to formulate HW-SW partitioning problems using the $\nu$Z tool.

@@

The mathematical modeling of the HW-SW partitioning problem was taken from~\cite{Arato2003,Mann2007}.

@@

The informal model can be described by five characteristics. First, there is only one software context, {\it i.e.}, there is just one general-purpose processor, and there is only one hardware context. The components of the system must be mapped to either one of these two contexts. Second, the software implementation of a component is associated with a software cost, which is the running time of the component. Third, the hardware implementation of a component has a hardware cost, which can be area, heat dissipation, and energy consumption. Fourth, based on the premise that hardware is significantly faster than software, the running time of the components in hardware is considered as zero. Finally, if two components are mapped to the same context, then there is no overhead of communication between them; otherwise, there is an overhead. The consequence of these assumptions is that scheduling does not need to be addressed in this work. Hardware components do not need scheduling, because the running time is assumed to be zero. Because there is only one processor, software components do not need to be scheduled as well. Therefore, the focus is only on the partitioning problem. That configuration describes a first-generation co-design, where the focus is on bipartitioning ~\cite{Teich2012}.

@@

\vspace{2 mm}
\noindent where $x$ represents the decision variable. Concerning the complexity of this problem, \cite{Arato2003} demonstrate that it is NP-Hard~\cite{Cormem}.

@@

The ILP and GA were taken from our previous studies~\cite{Trindade2015,Trindade2016}. Both use slack variables in order to be possible to represent constraints and to use commercial tools. However, GA had improvements from the parameters of related studies in order to increase the solution accuracy without producing timeout. The tuning was performed by empirical tests and resulted in changing of three parameters, which are passed to the function \textit{ga} of MATLAB~\cite{TheMathWorks2013}: the population size was set from $300$ to $500$, the Elite count changed from $2$ (default value) to $50$, and the number of generations changed from $100*NumberOfVariables$ (default) to $75$.

@@

Algorithm~\ref{Pseudocode-describing-ESBMC} shows ESBMC pseudocode with the same constraints and conditions placed on ILP and GA. Two values must be controlled to obtain the results and to perform the optimization. One is the initial software cost, as defined in Section~\ref{Formal-Model}. The other is the halting condition (code violation) that stops the algorithm.

@@

To optimize the CPU resources utilization without modifying the underlying SMT solver, the Open Multi-Processing (OpenMP) library~\cite{Dagum1998} is used in this present work as a front-end for ESBMC. Fig.~\ref{ESBMC-Multi-core} shows our first approach called ESBMC sequential-search ``ESBMC-SS''.

@@

Algorithm~\ref{Steps-Calculation-using-Intervals} describes how the customized binary search calculates and returns the step to be executed.

Note that Algorithm~\ref{Steps-Calculation-using-Intervals} is called from each worker in order to get the next step to execute if it exists; otherwise, either zero or a negative number is returned.
From lines $4$ to $9$, the algorithm finds the largest interval. Then, from line $10$ the largest interval is removed and the median is calculated in line $11$. After that, two new intervals are created, the left side (in line $14$) and the right side (in line $16$). At the end, the median is returned.

@@

\vspace {2 mm}
Algorithm~\ref{worker-sample} describes how the worker starts and monitors ESBMC instances. The algorithm starts by retrieving the step to be executed from the controller (line $1$), then initiates the ESBMC instance and obtains the process \textit{id} from the forked process (line $2$). While the step is being executed, the controller checks whether this step is still needed (line $4$). If not, then the ESBMC instance is killed (line $5$) and the worker is free to initiate another step.

@@

Algorithm~\ref{vZ-pseudocode} encodes the objective function and constraints related to the HW-SW partitioning problem using $\nu$Z functions~\cite{Bjorner2014}. A $\nu$Z logical context must firstly be created (line $2$), in order to add constraints and to check whether a given model exists to the set of constraints. Note that the number of nodes and edges, software, hardware, and communications costs as well as the incidence matrix E must also be declared.

@@

This section is split into three parts. The setup is described in Section~\ref{Experimental-Setup}, while Section~\ref{Benchmarks-Description} describes all benchmarks that were used for performing the experimental evaluation. Section~\ref{Experimental-Results} reports a comparison among Matlab~\cite{TheMathWorks2013}, ESBMC-SS, ESBMC-PS, ESBMC-PB, and ESBMC-$\nu$Z using a set of standard HW-SW partitioning benchmarks~\cite{Mann2007}.

@@

ESBMC v$2$.$0$ running on a $64$-bit Ubuntu $14$.$04$.$1$ LTS operating system was used. A parallel approach of the ESBMC-SS, ESBMC-PS, ESBMC-PB were implemented in C$++$11. Version $2$.$0$.$1$ of Boolector SMT-solver~\cite{Brummayer2009} (freely available) was used as the default solver for ESBMC. ESBMC-$\nu$Z as a built-in tool to Z3 was also used~\cite{Bjorner2014}.  For ILP and GA formulations, MATLAB R$2013$a from MathWorks with Parallel Computing Toolbox was used~\cite{TheMathWorks2013}. MATLAB is a dynamically typed high-level language, known as the state-of-the-art mathematical software~\cite{Tranquillo2011} and is widely used by the engineering community~\cite{Hong2010}.

@@

%----------------------------------------------
\subsection{Description of Benchmarks}
\label{Benchmarks-Description}
%----------------------------------------------
To perform the experiments, some benchmarks provided by~\cite{Mann2007} were used, as shown in Table~\ref{tab:Description-of-Benchmarks}. The nodes in the graphs correspond to high-level language instructions. Software and communication costs are time dimensional, and hardware costs represent the occupied area. The first three benchmarks are extracted from MiBench~\cite{Guthaus2001}. The clustering and fuzzy benchmarks are designed from~\cite{Mann2007} and are significantly large benchmarks. From the same authors, very complex benchmarks to test the limits of the applicability of techniques were used (RC$6$ and Mars).

\vspace{5 mm}
\hspace{-5 mm}
\begin{minipage}{\linewidth}
\centering
\small
\sffamily\small
\tabulinesep=4pt
\captionof{table}{Description of Benchmarks.} 
\begin{tabular}[c]{m{1.5cm}m{0.7cm}m{0.7cm}m{4.0 cm}}
\toprule[1.5pt]
\textbf{Name} & \textbf{Nodes} & \textbf{Edges} & \textbf{Description}\\
\midrule
\verb|CRC32| & \verb|25| & \verb|32| & \rmfamily 32-bit cyclic redundancy check ~\cite{Guthaus2001}\\
\hline
\verb|Patricia| & \verb|21| & \verb|48| & \rmfamily
\vspace{0.8 mm} Routine to insert values in Patricia Tree \cite{Guthaus2001}\\
\hline
\verb|Dijkstra| & \verb|26| & \verb|69| & \rmfamily \vspace{0.8 mm} Computer shortest paths in a graph ~\cite{Guthaus2001}\\
\hline
\verb|Clustering| & \verb|150| & \verb|331| & \rmfamily \vspace{0.8 mm} Image segmentation algorithm in a medical application\\
\hline
\verb|RC6| & \verb|329| & \verb|448| & \rmfamily \vspace{0.8 mm} RC6 cryptography graph algorithm\\
\hline
\verb|Fuzzy| & \verb|261| & \verb|422| & \rmfamily \vspace{0.8 mm} Clustering algorithm based on fuzzy logic\\
\hline
\verb|Mars| & \verb|417| & \verb|600| & \rmfamily 
\vspace{1 mm} MARS cipher from IBM algorithm\\
\bottomrule[1.5pt]
\label{tab:Description-of-Benchmarks} 
\end{tabular}\par
\bigskip
\end{minipage}

>>>>

\authorA{Author A}
\affA{University A\\
E-mail: \{nonono\}@nonono.com}

\authorB{Author B}
\affB{University B\\
E-mail: \{nonono\}@nonono.com}

%\authorA{Alessandro Trindade, Hussama Ismail}
%\affA{Federal University of Amazonas \\
%E-mail: \{alessandro.b.trindade, hussamaismail\}@gmail.com}

%\authorB{Edilson Galv\~ao, Renato Degelo}
%\affB{Federal University of Amazonas \\
%E-mail: \{rdegelo, esj.galvao\}@gmail.com}

%\authorC{Helder Silva, and Lucas Cordeiro}
%\affC{Federal University of Amazonas \\
%E-mail: prhsilva2012@gmail.com, lucascordeiro@ufam.edu.br}

\begin{abstract}
We present new alternative approaches to solve the hardware and software partitioning problem. First we use Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) in conjunction with a multi-core support using Open Multi-Processing to create four variants to solve the partitioning problem. In a nutshell, the multi-core SMT-based BMC approach allows initializing many verification instances based on the number of available processing cores. Each instance checks for a different optimum value until the optimization problem is satisfied. We implement our algorithms on top of the Efficient SMT-Based Context-Bounded Model Checker (ESBMC). Other approach integrate the maximum satisfiability solver $\nu$Z tool into ESBMC. $\nu$Z provides a portfolio of approaches for solving linear optimization problems over SMT formulas. We compare all our approaches to a state-of-the-art optimization tool (MATLAB). Experimental results show that there is no single optimization tool to solve all hardware-software partitioning benchmarks.
\end{abstract}

\KEYWORD{hardware-software co-design, hardware-software partitioning, optimization, model checking, multi-core, maximum satisfiability.}

\REF{\textcolor{red}{to this paper should be made as follows: No no no. `Multi-Core Model Checking and Maximum Satisfiability Applied to Hardware-Software Partitioning', {\it International Journal of Embedded Systems}, Vol. x, No. x, pp.xxx\textendash xxx.}}

%Trindade, A., Ismail, H., Galv\~ao, E., Degelo, R., Silva, H., and Cordeiro, L.%
\begin{bio}
\textcolor{red}{
no no no.\vs{9}}

\textcolor{red}{\noindent No no no.\vs{8}}

\textcolor{red}{\noindent No no no.}
\end{bio}

@@

Starting at the 1990s, intensive research was performed in HW-SW partitioning, and several approaches proposed, as shown in~\cite{Arato2003} and~\cite{Mann2007}. In~\cite{Trindade2015} and~\cite{Trindade2016} was shown that it is possible to use Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT), implemented in a tool called Efficient SMT-Based Context-Bounded Model Checker (ESBMC) as defined in~\cite{Cordeiro2012}, in order to perform HW-SW partitioning in embedded systems. In~\cite{Trindade2016} the original ESBMC algorithm was corrected and the right algorithm caused a loss in terms of performance, so a multi-core sequential approach was introduced in order to bring the performance back and turn it similar, in some cases, to conventional optimization tools that are used to perform hardware-software partitioning.

The present work extends those studies since there is a substantial improvement in terms of the SMT-based verification methods, which has been extended and exhauted with a multi-core parallel- and binary-search approaches, as well as, the integration of the Maximum SMT (MaxSMT) solver $\nu$Z into ESBMC, which is a state-of-art optimization tool based on SMT as defined in ~\cite{Bjorner2015}.

Here, we exploit the availability of multi-core processors and the tool OpenMP as defined in~\cite{Tang2009}; in particular, SMT-based verification methods are applied to the HW-SW partition problem in three different ways using a multi-core ESBMC approach with OpenMP: ESBMC-SS using a sequential-search (SS), ESBMC-PS using a parallel-search (PS), and ESBMC-PB using a binary-search (BS). Experimental results are compared to ILP (integral linear programming), GA (generic algorithms) in a multi-core version, and also to $\nu$Z, which supports only a single-core approach, according~\cite{Bjorner2015}. The ILP and GA algorithms are implemented with the optimization toolbox of Matlab, according~\cite{TheMathWorks2013}, while $\nu$Z is a built-in tool to the SMT solver Z3. All multi-core ESBMC approaches, together with $\nu$Z, are implemented with the ESBMC tool.

@@

Optimization is the act of obtaining the best result ({\it i.e.}, the optimal solution) under given circumstances as defined in~\cite{Rao2009}. There is no single method available for efficiently solving all optimization problems, according~\cite{Rao2009}. The most well-known technique is linear programming, which is an method applicable for the solution of problems in which the objective function and the constraints appear as linear functions of the decision variables. A particular case of linear programming is ILP, in which the variables can assume just integer values. Eq.~\ref{linear-programming-problem} shows a typical linear programming problem, where $A$ and $b$ are vectors or matrixes that describe the constraints

@@

\vspace{2 mm}
In some cases, the time to find a solution using ILP is impractical. Even with the use of powerful computers, a problem can take hours before an optimal solution is reached. If the optimization problem is complex, some heuristics can be used to solve the same problem faster, according~\cite{Rao2009}, {\it e.g.}, those used in the GA. The only drawback is that the found solution may not be the global minimum or maximum. Alternatively, tools such as ESBMC and $\nu$Z can be used to solve optimization problems so that the global minimum or maximum solution is found. The following sections describe the main features of ESBMC and $\nu$Z tools.

@@

Among the recent model checking techniques, there is one that combines model checking with satisﬁability solving. This technique, known as bounded model checking (BMC), does a very fast exploration of the state space, and for some types of problems, it offers large performance improvements over previous approaches, as shown in~\cite{Biere2009}. In particular, BMC based on Boolean Satisfiability (SAT) has been introduced as a complementary technique to binary decision diagrams for alleviating the state explosion problem, according~\cite{Clarke2001}.

The basic idea of BMC is to check the negation of a given property at a given depth: given a transition system $M$, a property $\phi$, and a bound $k$, BMC unrolls the system $k$ times and translates it into a verification condition (VC) $\psi$  such that $\psi$ is satisfiable if and only if $\phi$ has a counterexample of depth $k$ or less, as defined in~\cite{Biere2009}. To cope with increasing software complexity, SMT solvers can be used as back-ends for solving the generated VCs, as shown in~\cite{Cordeiro2012},~\cite{Armando2009},~\cite{Ganai2006}.

In this study, ESBMC has been used as a BMC tool to solve HW-SW partitioning problems, as shown in~\cite{Cordeiro2012}. 
In particular, there are two directives in ESBMC that can be used to guide it to solve an optimization problem: ASSUME and ASSERT. The directive ASSUME is responsible for ensuring the compliance of constraints (software costs), and the directive ASSERT controls the halt condition (minimum hardware cost). Then, with some C/C++ code, it is possible to guide ESBMC to solve optimization problems.

@@

Fig.~\ref{ESBMC-Architecture} shows the current ESBMC architecture, which consists of the C/C$++$ parser, GOTO Program, GOTO Symex, and SMT solver, according~\cite{Ramalho2013}. In particular, ESBMC compiles C/C$++$ code into equivalent GOTO-programs ({\it i.e.}, control-flow graphs) using a gcc-compliant style. GOTO-programs can then be processed by the symbolic execution engine, called GOTO Symex, where two recursive functions compute the constraints ($C$) and properties ($P$); finally, it generates two sets of equations ({\it i.e.},\:$C \land \neg P$), which are checked for satisfiability by an SMT solver.

The main factor for ESBMC to use only a single-core relies on its back-end ({\it i.e.}, SMT Solver). Currently, the SMT solvers supported by ESBMC are: Z3, as shown in~\cite{DeMoura2008}; Boolector, as shown in~\cite{Brummayer2009}; MathSAT, as shown in~\cite{Barrett2011}; CVC4, as shown in~\cite{Bozzano2005}; and Yices, as shown in~\cite{Dutertre2014}. Most of them do provide neither multi-threaded support nor a parallel version to solve the generated SMT equations.

@@

The OpenMP is a set of directives for parallel programming that augments C/C$++$ and Fortran languages, as defined in~\cite{OpenMP1998}. OpenMP supports most processor architectures and operating systems, {\it e.g.}, Solaris, AIX, HP-UX, Linux, Mac OS X, and Windows. OpenMP uses a portable and very robust model to facilitate the development of parallel applications for a variety of platforms.

In particular, OpenMP uses the \textit{fork-join} model of parallel execution, as shown in~\cite{OpenMP1998}. The main thread executes the sequential parts of the program; if a parallel region is encountered, then it forks a team of worker threads. After the parallel region finishes ({\it i.e.}, the API waits until all threads terminate), then the main procedure returns to the single-threaded execution mode, as shown in~\cite{Wu2014}.

@@

In this study, the SMT solver Z3 is used to check for the satisfiability of formulas generated from the HW-SW partitioning problem, as shown in~\cite{Bjorner2014}. In particular, we exploit the use of MaxSMT solver $\nu$Z, which is implemented on top of the SMT solver Z3, in order to solve optimization problems; $\nu$Z base function is to optimize objective functions, which formulate optimized criteria, within the logical context of constraints.~$\nu$Z also includes an incremental version of the Maximum Resiliency (MaxRes), as shown in~\cite{Federica2008}, in order to achieve Maximum Satisfiability (MaxSAT), as defined in~\cite{NarodytskaN} and a Simplex to solve numbers without defined patterns.

In $\nu$Z, MaxSAT is responsible for the restrictions, while OptSMT optimizes linear arithmetic objectives, as shown in~\cite{Bjorner2015}. In summary, $\nu$Z provides three main functions that extend Z3 for solving optimization problems, which are: \textit{maximize}, \textit{minimize}, and \textit{assert-soft}.

@@

\vspace{2 mm}
Fig.~\ref{vZ-Architecture} shows the $\nu$Z architecture. Initially, the SMT formula with objectives is converted to $0-1$ constraints, which leads to a Pseudo-Boolean Optimization (PBO), as shown in~\cite{Barth1995} and~\cite{Vasco2005}. If there are many objective functions, $\nu$Z invokes OptSAT for arithmetic or MaxSAT for soft constraints. For constraints using real values, $\nu$Z combines linear arithmetic objectives and uses only one instance of OptSMT. When ``soft constrains'' is used in the mode `` lexicographic'', $\nu$Z invokes MaxSAT using multiple calls for its engine.

@@

Z3 is available for platforms in C, C$++$, Java, .NET, and Python; it is possible to download Z3 with $\nu$Z from its github repository in~\cite{Z3API}. In this work, the python API is used to formulate HW-SW partitioning problems using the $\nu$Z tool.

@@

The mathematical modeling of the HW-SW partitioning problem was taken from~\cite{Arato2003} and~\cite{Mann2007}.

@@

The informal model can be described by five characteristics. First, there is only one software context, {\it i.e.}, there is just one general-purpose processor, and there is only one hardware context. The components of the system must be mapped to either one of these two contexts. Second, the software implementation of a component is associated with a software cost, which is the running time of the component. Third, the hardware implementation of a component has a hardware cost, which can be area, heat dissipation, and energy consumption. Fourth, based on the premise that hardware is significantly faster than software, the running time of the components in hardware is considered as zero. Finally, if two components are mapped to the same context, then there is no overhead of communication between them; otherwise, there is an overhead. The consequence of these assumptions is that scheduling does not need to be addressed in this work. Hardware components do not need scheduling, because the running time is assumed to be zero. Because there is only one processor, software components do not need to be scheduled as well. Therefore, the focus is only on the partitioning problem. That configuration describes a first-generation co-design, where the focus is on bipartitioning, as shown in~\cite{Teich2012}.

@@

\vspace{2 mm}
\noindent where $x$ represents the decision variable. Concerning the complexity of this problem, \cite{Arato2003} demonstrate that it is NP-Hard, as defined in~\cite{Cormem}.

@@

The ILP and GA were taken from our previous studies, from~\cite{Trindade2015} and~\cite{Trindade2016}. Both use slack variables in order to be possible to represent constraints and to use commercial tools. However, GA had improvements from the parameters of related studies in order to increase the solution accuracy without producing timeout. The tuning was performed by empirical tests and resulted in changing of three parameters, which are passed to the function \textit{ga} of MATLAB, as shown in~\cite{TheMathWorks2013}: the population size was set from $300$ to $500$, the Elite count changed from $2$ (default value) to $50$, and the number of generations changed from $100*NumberOfVariables$ (default) to $75$.

@@

Figure~\ref{Pseudocode-describing-ESBMC} shows ESBMC pseudocode with the same constraints and conditions placed on ILP and GA. Two values must be controlled to obtain the results and to perform the optimization. One is the initial software cost, as defined in Section~\ref{Formal-Model}. The other is the halting condition (code violation) that stops the algorithm.

@@

To optimize the CPU resources utilization without modifying the underlying SMT solver, the Open Multi-Processing (OpenMP) library, according~\cite{Dagum1998}, is used in this present work as a front-end for ESBMC. Fig.~\ref{ESBMC-Multi-core} shows our first approach called ESBMC sequential-search ``ESBMC-SS''.

@@

Algorithm at Figure~\ref{Steps-Calculation-using-Intervals} describes how the customized binary search calculates and returns the step to be executed.

Note that algorithm of Figure~\ref{Steps-Calculation-using-Intervals} is called from each worker in order to get the next step to execute if it exists; otherwise, either zero or a negative number is returned.
From lines $4$ to $9$, the algorithm finds the largest interval. Then, from line $10$ the largest interval is removed and the median is calculated in line $11$. After that, two new intervals are created, the left side (in line $14$) and the right side (in line $16$). At the end, the median is returned.

@@

\vspace {2 mm}
Algorithm of Figure~\ref{worker-sample} describes how the worker starts and monitors ESBMC instances. The algorithm starts by retrieving the step to be executed from the controller (line $1$), then initiates the ESBMC instance and obtains the process \textit{id} from the forked process (line $2$). While the step is being executed, the controller checks whether this step is still needed (line $4$). If not, then the ESBMC instance is killed (line $5$) and the worker is free to initiate another step.

@@

Algorithm of Figure~\ref{vZ-pseudocode} encodes the objective function and constraints related to the HW-SW partitioning problem using $\nu$Z functions, as seen in~\cite{Bjorner2014}. A $\nu$Z logical context must firstly be created (line $2$), in order to add constraints and to check whether a given model exists to the set of constraints. Note that the number of nodes and edges, software, hardware, and communications costs as well as the incidence matrix E must also be declared.

@@

This section is split into three parts. The setup is described in Section~\ref{Experimental-Setup}, while Section~\ref{Benchmarks-Description} describes all benchmarks that were used for performing the experimental evaluation. Section~\ref{Experimental-Results} reports a comparison among MATLAB, ESBMC-SS, ESBMC-PS, ESBMC-PB, and ESBMC-$\nu$Z using a set of standard HW-SW partitioning benchmarks, as shown in~\cite{Mann2007}.

@@

ESBMC v$2$.$0$ running on a $64$-bit Ubuntu $14$.$04$.$1$ LTS operating system was used. A parallel approach of the ESBMC-SS, ESBMC-PS, ESBMC-PB were implemented in C$++$11. Version $2$.$0$.$1$ of Boolector SMT-solver, as shown in~\cite{Brummayer2009} (freely available) was used as the default solver for ESBMC. ESBMC-$\nu$Z as a built-in tool to Z3 was also used, as shown in~\cite{Bjorner2014}.  For ILP and GA formulations, MATLAB R$2013$a from MathWorks with Parallel Computing Toolbox was used, as shown in~\cite{TheMathWorks2013}. MATLAB is a dynamically typed high-level language, known as the state-of-the-art mathematical software, as shown in~\cite{Tranquillo2011} and is widely used by the engineering community, as shown in~\cite{Hong2010}.

@@

%----------------------------------------------
\subsection{Description of Benchmarks}
\label{Benchmarks-Description}
%----------------------------------------------
To perform the experiments, some benchmarks provided by~\cite{Mann2007} were used, as shown in Table~\ref{tab:Description-of-Benchmarks}. The nodes in the graphs correspond to high-level language instructions. Software and communication costs are time dimensional, and hardware costs represent the occupied area. The first three benchmarks are extracted from MiBench, as shown in~\cite{Guthaus2001}. The clustering and fuzzy benchmarks are designed from~\cite{Mann2007} and are significantly large benchmarks. From the same authors, very complex benchmarks to test the limits of the applicability of techniques were used (RC$6$ and Mars).

\vspace{5 mm}
\hspace{-5 mm}
\begin{minipage}{\linewidth}
\centering
\small
\sffamily\small
\tabulinesep=4pt
\captionof{table}{Description of Benchmarks.} 
\begin{tabular}[c]{m{1.5cm}m{0.7cm}m{0.7cm}m{4.0 cm}}
\toprule[1.5pt]
\textbf{Name} & \textbf{Nodes} & \textbf{Edges} & \textbf{Description}\\
\midrule
\verb|CRC32| & \verb|25| & \verb|32| & \rmfamily 32-bit cyclic redundancy check, as shown in~\cite{Guthaus2001}\\
\hline
\verb|Patricia| & \verb|21| & \verb|48| & \rmfamily
\vspace{0.8 mm} Routine to insert values in Patricia Tree, as shown in ~\cite{Guthaus2001}\\
\hline
\verb|Dijkstra| & \verb|26| & \verb|69| & \rmfamily \vspace{0.8 mm} Computer shortest paths in a graph, as shown in~\cite{Guthaus2001}\\
\hline
\verb|Clustering| & \verb|150| & \verb|331| & \rmfamily \vspace{0.8 mm} Image segmentation algorithm in a medical application\\
\hline
\verb|RC6| & \verb|329| & \verb|448| & \rmfamily \vspace{0.8 mm} RC6 cryptography graph algorithm\\
\hline
\verb|Fuzzy| & \verb|261| & \verb|422| & \rmfamily \vspace{0.8 mm} Clustering algorithm based on fuzzy logic\\
\hline
\verb|Mars| & \verb|417| & \verb|600| & \rmfamily 
\vspace{1 mm} MARS cipher from IBM algorithm\\
\bottomrule[1.5pt]
\label{tab:Description-of-Benchmarks} 
\end{tabular}\par
\bigskip
\end{minipage}
